---
title: "Exam 2 Fiancial Econometrics"
author: "Juan David Torres"
date: "May 5 2021"
output:
  html_document:
    df_print: paged
---

# Problem 1
```{r}
library(readxl)
download.file("http://www.apradie.com/datos/dataex1.xlsx", "dataex1.xlsx", mode="wb")
data<-read_excel("dataex1.xlsx")
```
Calculations
```{r}
# Financial leverage
data$fl <- data$ltdebt / data$ta

# Earnings per share
data$eps <- data$ebit / data$sharesoutstanding

# Earnings per share deflated by price
data$epsp <- data$eps / data$adjustedstockprice

# Data in the quarter 2015-01-01
data$q<-as.Date(data$quarter)
data2015 <- data[data$q=="2015-01-01",]
```
Run the regression
```{r}
reg1 <- lm(data2015$F1r ~ data2015$epsp + data2015$fl, na.action = na.omit)
s_reg1 <- summary(reg1)
s_reg1
```

Calculation 95% confidence interval
```{r}
b1low = 0.52622 - 2*0.36710
b1high = 0.52622 + 2*0.36710
b1low
b1high

b2low = -0.26020 - 2*0.13410
b2high = -0.26020 + 2*0.13410
b2low
b2high
```

QUESTION: Provide a clear INTERPRETATION of the coefficients and their corresponding p-values.
ANSWER: 
CREATED A LINEAR REGRESSION MODEL FOR THE QUARTER 2015-01-01 TO EXAMINE WETHER THE STOCK RETURN IS AFFECTED BY THE FINANCIAL LEVERAGE AFTER CONSIDERING THE EFFECT OF THE EARNINGS PER SHARE DEFLATED BY PRICE. THE DEPENDENT VARIBLE IS THE STOCK RETURN AND THE INDEPENDENT VARIABLES ARE THE FINANCIAL LEVERAGE AND THE EARNINGS PER SHARE DEFLATED BY PRICE. 

AFTER CONSIDERING THE EFFECT OF THE FINANCIAL LEVERAGE THE EARNINGS PER SHARE DEFLATED BY PRICE WHEN THE EPSP INCREASES THE EFFECT IT HAS ON THE STOCK PRICE IS POSITIVE BUT NOT SIGNIFICANT BECAUSE THE P VALUE IS 0.1553 AND BECASE IS BIGGER THAN 0.05 THEN IT IS NOT SIGNIFICANT. IN AVERAGE WHEN THE EPSP WHEN IT INCREASES BY 1 UNIT THE STOCK RETURN WILL INCREASE BY 52.622% AFTER CONSIDERING THE EFFECT OF THE FINANCIAL LEVERAGE. THE 95% CONFIDENCE INTERVAL OF THE B1 COEFFICIENT IS -0.20798 AND 1.26042 WHICH MEANS THAT WITH A 95% OF CONFIDENCE THE BETA1 WILL VARY BETWEEN THOSE TWO VALUES.

AFTER CONSIDERING THE EFFECT OF THE EPSP THE FINANCIAL LEVERAGE HAS AN INVERSE RELATION WITH THE STOCK RETURN BUT NOT SIGNIFICANT BECAUSE THE P VALUE IS 0.0555 AND IT IS HIGHER THAN 0.05. WHEN THE FINANCIAL LEVERAGE INCREASES BY 1 UNIT THE STOCK RETURN DECREASES 26.02% IN AVERAGE AFTER CONSIDERING THE EFFECT OF THE EPSP. THE 95% CONFIDENCE INTERVAL OF B2 IS -0.5284 AND 0.008 WHICH MEANS THAT THE BETA2 VALUE CAN VARY FROM THOSE TWO VALUES WITH A 95% CONFIDENCE.

WHEN THE EPSP AND THE FL ARE BOTH 0 THEN THE STOCK RETURN IS IN AVERAGE EQUAL TO 0.00246.

THE FORMULA TO PREDICT WILL BE R = 0.00246 + 0.52622 * EPSP - 0.26020 * FL

# Problem 2
```{r}
table(data$sizetype)
```

Because I have already calculated the EPS on the first problem I wont calculate them again because it is not necessary.
Calculation for the dependent and  independent variables
```{r}
library(statar)

data$epsp_w <- winsorize(data$epsp,probs = c(0.02,0.98))

data$cfr = data$cashflowoper / data$ta
data$cfr_w = winsorize(data$cfr,probs = c(0.01,0.99))

# Categorical variable
data$size1 <- factor(data$sizetype, c("small","medium","big"), labels = c("Small", "Medium", "Big")) 

# Look at the coding
contrasts(data$size1)
```

Run the regression
```{r}
reg2 <- lm(data$epsp_w ~ data$cfr_w + data$size1, na.action = na.omit)
s_reg2 <- summary(reg2)
s_reg2
```

```{r}
b1low_2 = 0.115252 - 2*0.038837
b1high_2 = 0.115252 + 2*0.038837
b1low_2
b1high_2

```

QUESTION: Provide a clear INTERPRETATION of the coefficients and their corresponding p-values.
ANSWER:
I CREATED A LINEAR REGRESSION MODEL FOR ALL OF THE QUARTERS GIVEN IN THE DATA TABLE, THIS TO EXAMINE WETHER THE CASHFLOW GENERATED FROM THE OPERATIONS AND THE CATEGORICAL VARIABLE SIZE TYPE AFFECT THE EARNINGS PER SHARE DEFLATED BY PRICE AND WINSORIZED.

AFTER CONSIDERING THE EFFECT OF THE CATEGORICAL VARIABLE SIZETYPE THE CASH FLOW RATIO WINSORIZE HAS A POSITIVE AND SIGNIFICANT RELATIONSHIP WITH THE EPSP WINSORIZED BECAUSE THE P VALUE IS 0.00306 WHICH IS SMALLER THAN 0.05 WHICH MEANS IS SIGNIFICANT. AN INCREASE OF 1 UNIT IN THE CFR_W INCREASES THE EPSP_W BY 0.115252 CONSIDERING THE EFFECT OF THE SIZE TYPE. THE 95% CONFIDENCE INTERVAL OF THE CFR_W COEFFICIENT IS 0.037578 AND 0.192926 WHICH MEANS THAT THE COEFFICIENT IS POSITIVE AND THAT THE COEFFICIENT VARY FROM THOSE VALUE AT 95% CONFIDENCE.

AFTER CONSIDERING THE EFFECT OF THE CFR_W IT CAN BE SAID THAT IN COMPARISSON WITH THE SMALL SIZE THE MEDIUM SIZE FIRMS DECREASES -0.037932 THE EPSP_W VALUE AND THE DIFFERENCE IS STATISTICALLY SIGNIFICANT BECAUSE THE VALUE OF P IN THE MEDIUM SIZED IS 3.01 * e^05 AND BECAUS IT HIGHER THAN -0.05 IS SIGNIFICANT. THE VALUE OF THE MEDIUM COEFFICIENT CORRESPONDS TO THE VERTICAL DISTANCE BETWEEN THE MEDIUM SIZE AND THE SMALL SIZE. THE FORMULA FOR A MEDIUM SIZED FIRM WOULD BE EPSP_W = (0.054687 -0.037932)+ 0.115252 * CFR 

ON THE OTHER HAND, AFETR CONSIDERING THE EFFECT OF THE CFR_W THE BIG SIZE FIRMS DECREASE THE EPSP_W BY -0.001218 IN CONTRAST TO THE SMALL SIZED FIRMS, THIS DIFFERENCE IS NOT STATISTICALLY SIGNIFICANT AS THE P VALUE IS 0.77142 AND BECAUSE IT IS BIGGER THAN 0.05 THEN THE VERTICAL DISTANCE BETWEEN THE COEFFICIENT OF BIG SIZED COMPANIES AND SMALL FIRMS IS NOT SIGNIFICANT.
THE FORMULA FOR A BIG SIZED FIRM WOULD BE EPSP_W = (0.054687 - 0.001218)+ 0.115252 * CFR 

AND THE FORMULA FOR A SMALL SIZED FIRM WILL BE EPSP_W = 0.054687 + 0.115252 * CFR 

