---
title: "Workshop 3 Fiancial Econometrics"
author: "Juan David Torres"
date: "Feb 24 2021"
output:
  html_document:
    df_print: paged
---
# Review of R objects

In R any information is stored in objects. There are different data classes with its own data structure.
Data structures
1. Vectors
2. Matrix
3. Data frame
4. Lists

# Vectors
A vector can be of any data type.
Create a small collection of numbers
To create a vector we use the function c() which mean combine ande values separated with commas.

```{r}
stock_1 <- c(1.3, 1.5, 1.28)
stock_1

```

An integer vector creation in R with consecutive numbers

```{r}
stock_2 <- 1:10
stock_2
```

Arithmetic operations with numeric vector
```{r}
stock_3 <- stock_1 - 1
stock_3
```

Arithmetic operations with two or more vectors

```{r}
stock_4 <- stock_2 + stock_3
```

Sums all the stock_3 values to stock_2 consecutively as if stock 3 was of the same length than stock two
```{r}
stock_4
```

Atomic data classes

Numeric
```{r}
x <- c(2,3.43,4.21)
```

Character
```{r}
x <- c("2", "hi", "Econometrics I")
```

Integer
```{r}
x <- c(2L,3L,4L)
```

Logical (True/False)
```{r}
x <- c(TRUE,FALSE,FALSE,TRUE)
```

Complex (Imaginary numbers)
```{r}
z <- c(5i,4i)
```

Access a particular element of a vector. First create a vector from 101-105.
```{r}
vector1 <- seq(from = 101, to = 105)
vector1
```

Access the 3 element 
```{r}
vector1[3]
```


Access more than one element by its index we use a vector with the indexes we are looking to access
```{r}
vector1[c(1,3,5)]
```


To change a certain value of an index
```{r}
vector1[2] <- 109
vector1
```

Displays horizontal although they are considered as vertical
```{r}
vector2 <- seq(1,100)
vector2
```

# Matrix

Matrices are vectors with dimensions
Use the function matrix() specifying the number of rows with nrow and the number of columns with ncol.

```{r}
matrix(1:4, nrow = 2, ncol = 2)
```

Create a matrix joining 2 or more vectors with rbind (vectors as rows) or cbind(vectors as columns)

```{r}

stock1 <- c(1,2,3)
stock2 <- c(90,91,92)

matrix1 <-cbind(stock1,stock2)

matrix1
```

```{r}
matrix2 <-rbind(stock1,stock2)

matrix2
```

Access an element on a matrix with [] the first element indicates the row number and the second the column number.

Note: The indexes don't start with 0
```{r}
matrix1[1,2]
```

Refer to a whole column or a whole row by leaving empty the element you don't want to specify.
```{r}
matrix1[2,]
```

They have to be from the same class meaning numeric, string and so on.

# Data Frames
Data frames are used to store tabular data like excel. Data frames can store different classes of objects in different columns

To create a data frame we use data.frame()
```{r}
prices <- data.frame(Stocks = c("ALSEA","AMXL","FEMSA"), price = c(9.4, 5.5, 7.8))
prices
```

See the dimensions with dim(). Rows columns dimensions
```{r}
dim(prices)
```

Create data frame from an existing matrix with as.data.frame()
```{r}
my.matrix <- matrix(1:4, nrow = 2, ncol = 2)
my.df <- as.data.frame(my.matrix)
class(my.df)
```

The rbind and cbind can also be used in data frames
```{r}
prices.df <- prices
prices.df <- cbind(prices.df, behavior=c("up", "down", "up"))
prices.df
```


Data frames have attributes
```{r}
attributes(prices.df)
```

It has 3 attributes
You can change the name of the columns using colnames
```{r}
colnames(prices.df) <- c("ALSEA","AMXL","FEMSA")
attributes(prices.df)
```

Get the colnames and rownames
```{r}
colnames(prices.df)
rownames(prices.df)
```

Change the name of an specific column with the index
```{r}
colnames(prices.df)[3] <- "up-down"

colnames(prices.df)
```

# Creating an object from another object

Use the assignment operator <-

Add library quantmod to get information using getsymbols
```{r}
library(quantmod)
getSymbols("AAPL", from="2020-08-01", src="yahoo", periodicity="daily")

```

Last rows of AAPL
```{r}
tail(AAPL)
```


AAPL is an xts object, which is similar to a data frame but with a time/date index. 
Use $ to refer to an especific column

Getting the adjusted prices of AAPL
```{r}
AdPrice <- AAPL$AAPL.Adjusted
AdPrice
```

Get the natural logarithm of the adjusted prices
```{r}
lnAPPL <- log(AdPrice)
```

# Introduction to Hypothesis testing

Download monthly data of AAPL and MSFT
```{r}
library(quantmod)
getSymbols(c("AAPL", "MSFT"), from="2016-01-01", periodicity="monthly", src="yahoo")

```

Calculate the continuously compounded returns
```{r}
AAPL <- na.omit(AAPL)
MSFT <- na.omit(MSFT)
r_AAPL <- diff(log(Ad(AAPL)))
r_MSFT <- diff(log(Ad(MSFT)))
head(r_AAPL, 5)
```

```{r}
head(r_MSFT, 5)
```


# 3.1 T-test for AAPL

```{r}
# a) Null and alternative hypothesis
# H0: mean(r_AAPL) = 0
# Ha: mean(r_AAPL) > 0

# b)
# Calculation of the standard error
se_AAPL <- sd(r_AAPL, na.rm = TRUE) / sqrt(nrow(r_AAPL) - 1)
cat("Standard error r_AAPL" , se_AAPL,"\n")
```
```{r}
# c) 
# Calculation of the t-value. The 0 below is the H0
t_val_AAPL <- (mean(r_AAPL, na.rm=TRUE) - 0) / se_AAPL
cat("t-value AAPL = ", t_val_AAPL)
```
d) CONCLUSION OF THE T-TEST
AS THE T-VALUE IS 2.525702 THIS MEANS THAT THERE IS ENOUGH EVIDENCE AGAINST THE NULL HYPOTHESIS WHICH MEANS THAT THE NULL HYPOTHESIS IS REJECTED. 

WHEN THE T-VALUE IS BETWEEN -2 AND 2 IT MEANS THE NULL HYPOTHESIS CANNOT BE REJECTED BUT BECAUSE THE VALUE IS GREATER THAN 2 THEN IT CAN BE REJECTED. 
THIS MEANS THAT THE AAPL MEAN CC RETURNS ARE DIFFERENT FROM 0.


# 4.1 Running the T-test function from R
```{r}
ttest_AAPL <- t.test(as.numeric(r_AAPL), alternative = "greater")
print(ttest_AAPL$statistic)
```

THE T-VALUE IS THE SAME AS THE ONE CALCULATED ABOVE SO THIS MEANS THAT THE NULL HYPOTHESIS CAN BE REJECTED AND THE CALCULATIONS WERE CORRECTLY DONE.

# 3.2 T-test for MSFT

```{r}
# a) NULL AND ALTERNATIVE HYPOTHESIS
# H0: mean(r_MSFT) = 0
# Ha: mean(r_MSFT) > 0

# b) Calculation of the standard error
se_MSFT <- sd(r_MSFT, na.rm = TRUE) / sqrt(nrow(r_MSFT) - 1)
cat("Standard error r_MSFT" , se_MSFT,"\n")

```

```{r}
# c) Calculate the t-statistic
t_val_MSFT <- (mean(r_MSFT, na.rm=TRUE) - 0) / se_MSFT
cat("t-value AAPL = ", t_val_MSFT)
```

d) Conclusions of the t-test
SINCE THE T VALUE IS GREATER THAN 2 THIS MEANS THAT THE NULL HYPOTHESIS CAN BE REJECTED WHICH MEANS THAT THE MEAN CC RETURNS FROM MICROSOFT ARE GREATER THAN 0 WHICH IS THE ALTERNATIVE HYPOTHESIS, ALTHOUGH IT CAN ALSO BE THAT THE CC RETURNS BE LOWER THAN 0 BECAUSE IT IS ANOTHER ALTERNATIVE HYPOTHESIS.
SINCE IT IS NEEDED FOR THE T VALUE TO BE BETWEEN -2 AND 2 TO NOT REJECT THE NULL HYPOTHESIS AND THIS DIDN'T HAPPEND THEN WE CAN SAY THAT THE NULL HYPOTHESIS IS REJECTED.

# 4.2 Calculation of t-test using the t.test function from R
```{r}
ttest_MSFT <- t.test(as.numeric(r_MSFT), alternative = "greater")
print(ttest_MSFT$statistic)
```

IT GAVE THE SAME ANSWER WHICH MEANS THAT THE CALCULATIONS WERE DONE CORRECTLY.

# 5. T-test to compare both stocks returns

Calculations
```{r}
# Null hypothesis and alternative hypothesis
# H0: mean(r_AAPL) - mean(r_MSFT) = 0
# Ha: mean(r_AAPL) - mean(r_MSFT) <> 0

# Calculations of the t_value
N <- nrow(r_AAPL)-1
se_both <- sqrt( (1/N) * (var(r_AAPL, na.rm=TRUE) + var(r_MSFT, na.rm=TRUE)))
t_value_both <- (mean(r_AAPL, na.rm=TRUE) - mean(r_MSFT, na.rm=TRUE) - 0) / se_both
cat("t-value = ", t_value_both, "\n")
```

Using the t.test function
```{r}
ttest <- t.test(as.numeric(r_AAPL), as.numeric(r_MSFT), paired = FALSE, var.equal = FALSE)
names(ttest)
```

Get the t_value with the statistic function
```{r}
cat("t-vale from t.test =", ttest$statistic,"\n")
```

Get the p value with the p.value column
```{r}
cat("p-value = ", ttest$p.value)
```

SICNE THE T-VALUE IS BETWEEN -2 AND 2 THEN WE CANNOT REJECT THE HYPOTHESIS FUNCTION, MEANING THAT THERE IS NO DIFFERENCE BETWEEN THE AVERAGE CC RETURNS FROM AAPL AND MSFT, THEY ARE STATISTICALLY EQUAL. THE P_VALUE IS GREATER THAN 0.05, SINCE THE SMALLER THE VALUE OF P THE STRONGER THERE IS EVIDENCE TO REJECT H0, WHICH MEANS THAT THE NULL FUNCTION CAN'T BE REJECTED.

